### 2020-04-08 1_introduction.tex
Tweak text to save space.

One approach to comparing the performance of different systems is through \emph{proxy-} or
\emph{mini-apps}. Because they distill the key computational characteristics of a larger
application, mini-apps can provide insight
without the expense of developing a production code. However, despite the name, our experience is that
mini-apps still require significant investment to develop
to the level of quality needed for useful benchmarking. In many cases,
the effort to tune for performance exceeds the effort to develop a correct implementation. As a result, implementations of mini-apps
often do not reach the level of maturity required to compare
systems. Few published studies compare more than a handful of systems~\cite{LULESH13}.

@@

Adding a system to Task Bench involves implementing a set of standard
services, such as executing a task or data transfer. Note that, though
the benchmarks are described in terms of task graphs, this is simply a
convenient representation of the computation, and the underlying
system need not provide any native support for tasks. We provide
Task Bench implementations in systems as diverse as MPI and
Spark. Task Bench provides a core API that encapsulates functionality
shared among systems, which reduces implementation effort and makes it
much easier to achieve truly apples-to-apples comparisons between
systems.

This approach has allowed us to benchmark 15 very different parallel
and distributed programming systems (listed in
Table~\ref{tab:systems}).  By running all systems on common benchmarks
we were able to quantify phenomena
that have never before been measured.
Most strikingly, the overheads of the systems we examine vary by more than five orders
of magnitude, with popular, widely used systems at both ends of the spectrum!  Clearly
the slower systems have ``good enough'' performance for some class of applications, while presumably
providing advantages in programmer productivity.

@@

METG captures the essence of what is important in a
weak or strong scaling study, the behavior at the limit of
scalability. For weak scaling, METG(50\%) corresponds to the
smallest problem size that can be weak-scaled with 50\%
efficiency. For strong scaling, METG(50\%) can be used to compute the
scale at which performance can be expected to dip below 50\%
efficiency.  We note that METG(50\%) for a given runtime system will
vary with the particular application and the underlying hardware---i.e., METG(50\%)
is not a constant for a given system, but we will see that individual systems have
a characteristic range of METG(50\%) values and that there is additional insight
in the reasons that METG can vary.

>>>>

One approach to comparing performance is through \emph{proxy-} or
\emph{mini-apps}. Because they distill key computational characteristics of larger
applications, mini-apps offer insight
without the expense of developing production codes. However, despite the name, our experience is that
mini-apps still require significant investment to develop
to the level of quality needed for useful benchmarking. In many cases,
the effort to tune for performance exceeds the effort to develop a correct implementation. As a result, implementations of mini-apps
often do not reach the level of maturity required to compare
systems. Few published studies compare more than a handful of systems~\cite{LULESH13}.

@@

Adding a system to Task Bench involves implementing a set of standard
services, such as executing a task or data transfer. Though
benchmarks are described in terms of task graphs, this is simply a
convenient representation of the computation, and the underlying
system need not provide any native support for tasks. We provide
Task Bench implementations in systems as diverse as MPI and
Spark. Task Bench provides a core API that encapsulates functionality
shared among systems, which reduces implementation effort and makes it
much easier to achieve truly apples-to-apples comparisons between
systems.

This approach has allowed us to benchmark 15 very different parallel
and distributed programming systems (see
Table~\ref{tab:systems}).  By running all systems on common benchmarks
we were able to quantify phenomena
that have never before been measured.
Most strikingly, the overheads of systems we examine vary by more than five orders
of magnitude, with popular, widely used systems at both ends of the spectrum!  Clearly,
slower systems have ``good enough'' performance for some applications, while presumably
providing advantages in programmer productivity.

@@

METG captures the important essence of a
weak or strong scaling study, the behavior at the limit of
scalability. For weak scaling, METG(50\%) corresponds to the
smallest problem size that can be weak-scaled with 50\%
efficiency. For strong scaling, METG(50\%) can be used to compute the
scale at which efficiency can be expected to dip below 50\%.
We note that METG(50\%) for a given runtime system will
vary with the application and the underlying hardware---i.e., METG(50\%)
is not a constant for a given system, but we find that systems have
a characteristic range of METG(50\%) values and that there is additional insight
in the reasons that METG can vary.

### 2020-04-08 2_task_bench.tex
Tweaks for spacing and wording.
Tweak text to save space.

To explore as broad a space of application scenarios as possible, Task
Bench provides a large number of configuration parameters. The most
important parameters are described in
Table~\ref{tab:parameters}. These parameters control the size and
structure of the task graph, the type and duration of the kernels
associated with each task, and the amount of data associated with
each dependence edge in the graph.

@@

Dependencies between tasks are controlled by a configurable dependence
relation. The
dependence relation enumerates the set of tasks from the
previous time step each task depends on, permitting a wide variety
of patterns to be implemented that are relevant to real
applications: stencils,
sweeps, FFTs, trees, etc. Dependence relations may be
parameterized, such as picking the $K$ nearest neighbors, or $K$
distant neighbors. They may also vary over time, such as in the FFT pattern. The set of dependence relations is extensible, making it easy to add patterns to represent new classes applications. Table~\ref{tab:equations} shows equations for the
dependence relations of the patterns in Figure~\ref{fig:task-graphs},
where $t$ is timestep, $i$ is column, and $W$ is the width of the task
graph.

@@

@@

In addition to specifying the shape and structure of the task graph, the
Task Bench core API also provides implementations of the kernels
executed by each task as well as other utility routines (to parse
inputs and display results). A sample of the core API compute kernel
is shown in Listing~\ref{lst:compute-kernel}. In addition to reducing
the effort required to implement Task Bench, providing central
implementations of these services ensures that all Task Bench
implementations can be scripted uniformly and eliminates a potential
source of performance disparity that can be a pitfall for mini-apps.

>>>>

To explore as broad a space of application scenarios as possible, Task
Bench provides a large number of configuration parameters. These
parameters are described in
Table~\ref{tab:parameters}, and control the size and
structure of the task graph, the type and duration of kernels
associated with each task, and the amount of data associated with
each dependence edge in the graph.

@@

Dependencies between tasks are determined by a dependence
relation. The
dependence relation identifies the tasks from the
previous time step each task depends on, permitting a wide variety
of patterns to be implemented that are relevant to real
applications: stencils,
sweeps, FFTs, trees, etc. Dependence relations may be
parameterized, such as picking the $K$ nearest neighbors, or $K$
distant neighbors. They may also vary over time, such as in the FFT pattern. The set of dependence relations is extensible, making it easy to add patterns to represent new classes applications. Table~\ref{tab:equations} shows equations for the
dependence relations of the patterns in Figure~\ref{fig:task-graphs},
where $t$ is timestep, $i$ is column, and $W$ is the width of the task
graph.

@@

In addition to specifying the shape of the task graph, the
core API also provides implementations of the kernels
executed by each task as well as other utility routines (to parse
inputs and display results). A sample of the core API compute kernel
is shown in Listing~\ref{lst:compute-kernel}. In addition to reducing
the effort required to implement Task Bench, providing central
implementations of these services ensures that all Task Bench
implementations can be scripted uniformly and eliminates a potential
source of performance disparity that can be a pitfall for mini-apps.

@@

### 2020-04-08 3_implementation.tex
Apply Alex's comments.
Tweak text to save space.

Among distributed task-based systems, there can be a scalability
bottleneck due to enumerating tasks sequentially. PaRSEC and StarPU
resolve this by permitting the user to omit tasks that are not
directly connected in the dependence graph. The program is replicated
(for efficiency of the analysis), but each task is executed only
once. Because the dynamic checks to see if a task should be executed
are not free of cost, we also provide a PaRSEC implementation (under
the label \lstinline{shard}) which has been hand-written to avoid all
such checks. Regent performs an equivalent optimization at compile
time~\cite{ControlReplication17} that does not require user
intervention.

>>>>

Among distributed task-based systems, there can be a scalability
bottleneck due to enumerating tasks sequentially. PaRSEC and StarPU
allow users to omit tasks that are not
directly connected in the dependence graph. The program is replicated
(for efficiency of analysis), but each task is executed only
once. Because the dynamic checks to see if a task should be executed
are not free of cost, we also provide a PaRSEC implementation (labeled \lstinline{shard}) hand-written to avoid all
such checks. Regent performs an equivalent optimization at compile
time~\cite{ControlReplication17} that does not require user
intervention.

### 2020-04-08 4_metg.tex
Tweak text to save space.

Figures~\ref{fig:flops-mpi} and \ref{fig:efficiency-mpi} show how METG is
calculated. The application, in this case an MPI implementation of the
Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
run on a single Haswell node of Cori with a problem size large enough
that runtime is dominated by the application's inner loops. This result confirms
that the application is properly configured and that the
efficiency metric is achievable. The problem
size is then repeatedly reduced while maintaining exactly the same hardware and software
configuration (in particular, the same number of nodes and tasks). The
expectation is that as problem size shrinks,
performance will begin to drop and eventually approach zero, as shown in Figure~\ref{fig:flops-mpi}. Systems
with lower runtime overheads maintain higher performance at smaller
problem sizes compared to systems with higher overheads.

@@

METG(50\%) is the intersection of this curve with 50\% efficiency, as
shown by the red, dashed lines in Figures~\ref{fig:flops-mpi} and
\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
task granularity of
4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s in this
configuration. %% We use 50\% because that is generally an acceptable level
%% of efficiency in practice, and values above 90\% can misrepresent the
%% performance of some systems (see
%% Section~\ref{subsec:peak-performance-and-efficiency}).

METG has a well-defined
relationship with quantities of interest to application developers,
namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
\ref{fig:strong-scaling-mpi} show the weak and strong scaling behavior of the MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
figures, the vertical axis is shown as wall time to emphasize the
relationship to time-to-solution, but it could equivalently be shown
as task granularity (as the number of tasks per execution is
fixed). Intuitively, at
larger problem sizes MPI is perfectly efficient. This can be seen at
the top of each figure, with flat lines when weak scaling and
ideally-sloped downward lines when strong scaling. Inefficiency begins
to appear at smaller problem sizes, towards the bottom of the graph,
where lines become more compressed. At the
very bottom, the lines compress together as running time becomes
dominated by overhead. Note that the
shapes at the bottom of the strong and weak scaling curves are identical,
and conform to the shape of the METG curve (marked by the red, dashed line).

>>>>

Figure~\ref{fig:flops-mpi} shows how METG is
measured. We run the application (MPI Task Bench) on a Cori Haswell node
with a problem size large enough that runtime is dominated by kernel
execution. This result confirms
that the application is properly configured and that the
efficiency metric is achievable. The problem
size is then repeatedly reduced while maintaining exactly the same hardware and software
configuration (in particular, the same number of nodes and tasks). The
expectation is that as problem size shrinks,
performance will begin to drop and eventually approach zero. Systems
with lower runtime overheads maintain higher performance at smaller
problem sizes compared to systems with higher overheads.

@@

METG(50\%) is the intersection of the curve at 50\% efficiency, as
shown by the red, dashed lines in Figures~\ref{fig:flops-mpi} and
\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
task granularity of
4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s in this
configuration. %% We use 50\% because that is generally an acceptable level
%% of efficiency in practice, and values above 90\% can misrepresent the
%% performance of some systems (see
%% Section~\ref{subsec:peak-performance-and-efficiency}).

METG has a well-defined
relationship with quantities of interest such as
weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
\ref{fig:strong-scaling-mpi} show the weak and strong scaling of MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
figures, the vertical axis is shown as wall time to emphasize the
relationship to time-to-solution, but it could equivalently be shown
as task granularity (as the number of tasks per execution is
fixed). Intuitively, at
larger problem sizes MPI is perfectly efficient. This can be seen at
the top of each figure, with flat lines when weak scaling and
ideally-sloped downward lines when strong scaling. Inefficiency begins
to appear at smaller problem sizes, towards the bottom of the graph,
where lines become more compressed. At the
very bottom, the lines compress together as running time becomes
dominated by overhead. Note that the countour of the bottom of each
graph is identical and conforms to the METG curve (marked by the red,
dashed line).

### 2020-04-08 5_evaluation.tex
Tweak text to save space.

Most systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOP/s compared
to systems that share all cores between the application and runtime. Some of the
higher overhead systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this case ran for over 6 hours.

Figure~\ref{fig:efficiency} plots efficiency (as a percentage of
peak FLOP/s) against task granularity. As described in Section~\ref{sec:metg}, this plot is
used to calculate METG(50\%). The
red, dashed line shows the point where 50\% efficiency is achieved.
In most cases, task granularity asymptotes prior to the 50\% efficiency point,
though some systems continue to improve as efficiency drops further. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold instead of
measuring task scheduling throughput of empty tasks
(effectively METG(0\%)). Measuring performance using empty tasks can
reward implementation strategies, such as devoting nearly 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

The least complicated pattern (stencil) is most favorable
to MPI, as it provides no
opportunity for task parallelism. For the stencil pattern, the
dominating factor is the basic overhead of executing a task, which is
minimal for MPI as the Task Bench implementation simply executes tasks one after another in alternation with
communication phases. In contrast, the asynchronous execution
mechanisms of other systems are pure overhead in this scenario.
The gap between MPI and other systems shrinks as the complexity of the
communication pattern grows, and even reverses as task parallelism is
added in the form of multiple task graphs.

@@

Implicitly parallel systems such as PaRSEC, StarPU and Regent that
rely on runtime analysis to build the DAG can suffer from
scalability bottlenecks if every node must consider the tasks
executing on all other nodes. PaRSEC DTD and StarPU mitigate this
partially by allowing the user to omit tasks not directly
dependent on those executed by a given node. However, this DAG
trimming approach requires dynamic checks that scale with
the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
retains dynamic checks that limit scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
uses a compile-time optimization to improve
scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible. Clearly, choosing a representative dependence
pattern is important when estimating the performance of a workload or
class of workloads.

>>>>

Most systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
use; these systems take a minor hit in peak FLOP/s compared
to systems that share all cores between the application and runtime. Some of the
higher-overhead systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this case ran for over 6 hours.

Figure~\ref{fig:efficiency} plots efficiency (as a percentage of
peak FLOP/s) vs. task granularity. As described in Section~\ref{sec:metg}, this is
used to calculate METG(50\%). The
red, dashed line indicates 50\% efficiency.
In most cases, task granularity asymptotes prior to this point,
though some systems continue to improve at lower values. Accounting for this effect is one of the main arguments in
favor of using reasonable efficiency thresholds for METG instead of
empty tasks
(i.e., METG(0\%)). Empty tasks
reward strategies, such as devoting 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

The least complicated pattern (stencil) is most favorable
to MPI, as it provides no
opportunity for task parallelism. The
dominating factor in this case is the overhead of executing a task, which is
minimal for MPI as the Task Bench implementation simply executes tasks in alternation with
communication phases. The asynchronous
mechanisms of other systems are pure overhead in this scenario.
The gap between MPI and other systems shrinks as complexity grows, and even reverses as task parallelism is
added in the form of multiple task graphs.

@@

PaRSEC, StarPU and Regent
rely on runtime analysis that can suffer from
scalability bottlenecks if every node must consider the tasks
executing on all other nodes. PaRSEC DTD and StarPU mitigate this
partially by allowing the user to omit tasks not directly
dependent on those executed by a given node. However, this DAG
trimming approach requires dynamic checks that scale with
the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
retains dynamic checks that limit scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
uses a compile-time optimization to improve
scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels. Clearly, choosing a representative dependence
pattern is important when estimating the performance of a workload or
class of workloads.
